Title here

### 2020-05-17 paper/paper.Rmd
JCGS format and add some general optim difficulties

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Sherry Zhang 
    url: https://github.com/huizezhang-sherry
    affiliation: Monash University
    affiliation_url: https://www.monash.edu/
date: "`r Sys.Date()`"
output: distill::distill_article
bibliography: biblio.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
library(tidyverse)
library(ferrn)
library(patchwork)
library(ggrepel)

load(here::here("data/temp.rda"))
load(here::here("data/data_mult.rda"))
load(here::here("data/holes_1d_geo.rda"))
load(here::here("data/holes_2d_geo.rda"))
load(here::here("data/holes_2d_better.rda"))
load(here::here("data/holes_2d_better_max_tries.rda"))
load(here::here("data/holes_2d_better_random.rda"))
load(here::here("data/holes_2d_geo_polish.rda"))
load(here::here("data/holes_2d_better_polish.rda"))
load(here::here("data/holes_2d_better_max_tries_polish.rda"))
load(here::here("data/holes_2d_better_random_polish.rda"))
load(here::here("data/interrupt_yes.rda"))
load(here::here("data/interrupt_no.rda"))
```

@@

Tour provides a way to explore multivariate data interactively via an established tour path. A tour path is formed by interpolating between randomly generated plane. Different types of tours are available depends on the purpose of the exploration i.e. a grand tour is suitable for randomly exploring the data from different angles; a guided tour detects a particular structure in the data. a manual tour allows user to manually control the projection @cook1997manual.

@@

Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-diemsion projection of multivariate data. A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine. The projection pursuit index measures the "interestingness" of data defined as its departure from normality. Numerous indices have been proposed in the literature, including lengendre index @friedman1974projection, hermite index @hall1989polynomial, natural hermite index @cook1993projection, chi-square index @posse1995projection, LDA index @lee2005projection and PDA index @lee2010projection. An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index. The discussion of existing optimisation procedure will be discussed in the next section.

Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path. An illustration modified from @buja2005computational entails the geodesic interpolation bewteen the plane generated by projection pursuit.

![](figures/tour_path.svg)

## Optimisation methods existing in tourr

Because of the fact that projection pursuit index can be non-differentiable and that the projection basis is constrained to be orthonormal, the optimisation need to be a derivative-free, constrained procedure. Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data. Posse @posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis. The neighbourhood is defined via the radius of the p-dimenionsal sphere, $c$. The new basis is taken as the target basis if it has higher index value, or the sampling continues. If no basis is found to have higher index value after a certain humber of tries $n$, the radius $c$ is halved. The algorithm stops when the maximum numbere of iteration is attained or the radius $c$ is less than a pre-determined number. [Pursuit package uses this method and it works greate! But I don't think we implement this - although don't think it is too hard to do it].

Cook @cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable. Since some indices could be non-differentiable, the computation of derivative is replaced by a psudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood. Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.

Simulated annealing @bertsimas1993simulated @kirkpatrick1983optimization is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$. Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.

## problems experienced with existing versions

- the simulated annealing algorithm doesn't include the acceptance probability

- is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995). Any tools available to  better understand its searching process?

- it is not understood how the index value changes during interpolation - is any interpolating basis better candidate for the next-round current basis than the target basis?

- How does the projected basis found by the optim look like? Are they close to each other/ far away? Any high-d or low-d vis of the projection basis?

- Do I want to introduce search_polish in this article?

@@

Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) @wickham2011tourrpackage. A tour includes two major components: a *generator* that generating the projection basis according to projection pursuit and an *interpolator* that performing geodesic interpolation between the projection basis.

@@

@@

```{r echo = TRUE}
holes_1d_geo %>% explore_proj_pca(animate = TRUE)
```

>>>>

---
title: Title here

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Author 1
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of YYY, University of XXX

- name: Author 2
  affiliation: Department of ZZZ, University of WWW

keywords:
- 3 to 6 keywords
- that do not appear in the title

abstract: |
  1. check consistency of using PP for projection pursuit and PPI for projection pursuit index

bibliography: biblio.bib
output: rticles::asa_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

library(tidyverse)
library(ferrn)
library(patchwork)
library(ggrepel)

load(here::here("data/temp.rda"))
load(here::here("data/data_mult.rda"))
load(here::here("data/holes_1d_geo.rda"))
load(here::here("data/holes_2d_geo.rda"))
load(here::here("data/holes_2d_better.rda"))
load(here::here("data/holes_2d_better_max_tries.rda"))
load(here::here("data/holes_2d_better_random.rda"))
load(here::here("data/holes_2d_geo_polish.rda"))
load(here::here("data/holes_2d_better_polish.rda"))
load(here::here("data/holes_2d_better_max_tries_polish.rda"))
load(here::here("data/holes_2d_better_random_polish.rda"))
load(here::here("data/interrupt_yes.rda"))
load(here::here("data/interrupt_no.rda"))

```

@@

Tour provides a way to explore multivariate data interactively via an established tour path. A tour path is formed by interpolating between randomly generated plane. Different types of tours are available depends on the purpose of the exploration i.e. a grand tour is suitable for randomly exploring the data from different angles; a guided tour detects a particular structure in the data. a manual tour allows user to manually control the projection [@cook1997manual].

@@

Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-diemsion projection of multivariate data. A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine. The projection pursuit index measures the "interestingness" of data defined as its departure from normality. Numerous indices have been proposed in the literature, including lengendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection]. An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index. The discussion of existing optimisation procedure will be discussed in the next section.

Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path. An illustration modified from [@buja2005computational] entails the geodesic interpolation bewteen the plane generated by projection pursuit.

```{r}
knitr::include_graphics(path = here::here("figures/tour_path_new.png"))
```

## Optimisation methods in projection pursuit literature

As @friedman1974projection said "..., the technique use for maximizing the (one- and two-dimensional) projection index strongly influences both the statistical and the com- putational aspects of the procedure." The quality of the optimisation procedure largely affect the tour view and thus, the interesting projection one could possibly observe. An ideal optimisation procedure needs to have following characteristics:

- *Being able to handle non-differentiable index function*: the index function could be noisy and non-differentible

- *Being able to optimise with constraints*: the projection matrix is restricted to an orthornormal matrix.

- *Being able to reveal both local and global maximum*: Although the primary interest is to find the global maximum, distinct local structures are also of our interest.

Below present three existing methods in tour.

@posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis. The neighbourhood is defined via the radius of the p-dimenionsal sphere, $c$. The new basis is taken as the target basis if it has higher index value, or the sampling continues. If no basis is found to have higher index value after a certain humber of tries $n$, the radius $c$ is halved. The algorithm stops when the maximum numbere of iteration is attained or the radius $c$ is less than a pre-determined number. [Pursuit package uses this method and it works greate! But I don't think we implement this - although don't think it is too hard to do it].

@cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable. Since some indices could be non-differentiable, the computation of derivative is replaced by a psudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood. Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$. Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.

<!-- Because of the fact that projection pursuit index can be non-differentiable and that the projection basis is constrained to be orthonormal, the optimisation need to be a derivative-free, constrained procedure. Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data. -->

## problems and difficulties in PP optimisation

Below listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific for PP optimisation.

- *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large. Also, there are discussions on how to avoid getting trapped in a local optimal in the literature.

- *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible. Stochastic optimisation method could be an alternative to solve these problems.

- *computation speed*: The optimisation procedure needs to be fast to compute since tours produces real-time animation of the projected data.

- *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm producing the same result for every run. This practice supports reproducibility, while less efforts has been made to guarantee different seeds will provide the same result.

- *high-dimensional decision variable*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional. Researcher would be better off if they can understand the relative position of different projection matrix in the high-dimensional space.

- *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximises the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users. It would be interesting to see if the interpolated bases could, in reverse, help the optimisation reach faster convergence.

<!-- the interpolation can be seen as giving information on the direction of the derivative. -->

<!-- - is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995). Any tools available to  better understand its searching process? -->

*Think about how does your pacakge help people to understand optimisation*

 - diagnostic on stochastic optim
 - vis the prograssion of multi-parameter decision variable 
 - understanding learning rate - neightbourhood parameter
 - understand where the local & global maximum is found - trace plot - see if noisy function

@@

Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage]. A tour includes two major components: a *generator* that generating the projection basis according to projection pursuit and an *interpolator* that performing geodesic interpolation between the projection basis.

@@

## add something about general vis

@@

```{r echo = TRUE, eval = FALSE}
holes_1d_geo %>% explore_proj_pca(animate = TRUE)
```
